---
title: "HW 1"
subtitle: "Data 180, Professor Kennedy"
author: 
  name: "Sanskriti Rijal"
  email: "rijals@dickinson.edu"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output: 
  html_document
---
Make sure that you have accepted the hw invitation link, and working with your clone repository.

**Due date:** September 21 beginning of class.

**Reminder:** You can insert R code in your Rmd file as a chunk in the following way:
```{r}
# Your code goes here
```
Make sure to insert a chunk for each (sub)section of the assignment.

## Preliminary Task
Go to [datacamp.com](https://datacamp.com) and complete the tutorial "Introduction to R".


## Task 1
a) Store the data below in a data frame object named `my_df`.


|Person | Preference|
|:------|:----------|
|Alice | Twitter |
| Bob | Facebook |
| Charlie | Twitter |
| Denise | Instagram |
| Erin | Facebook |
| Frank | Twitter |
| Gisele | Twitter |
| Harry | Instagram |

```{r}
my_df <- data.frame(
  Person = c("Alice", "Bob", "Charlie", "Denise", "Erin", "Frank", "Gisele", "Harry"),
  Preference = c("Twitter", "Facebook", "Twitter", "Instagram", "Facebook", "Twitter", "Twitter", "Instagram")
)
print(my_df)
```


b) What are the dimensions of `my_df`?

```{r}
dimensions <- dim(my_df)
print(dimensions)
```


c) Write R code that counts the total number of users.

```{r}
totalUser <- nrow(my_df)
print(totalUser)
```

d) Write R code that counts the total number of users who prefer Twitter.
```{r}
userTwitter <- sum(my_df$Preference == "Twitter")
print(userTwitter)
```

e) Write R code that returns what Gisele prefers.
```{r}
gisele_prefer <- my_df$Preference[my_df$Person == "Gisele"]
print(gisele_prefer)
```
## Task 2
a) Create a vector named `my_vec` that includes all integers from 1 to 10 (intclusive).
```{r}
my_vec <- 1:10
print(my_vec)
```

b) Replace the second element in `my_vec` with the number `99`.
```{r}
my_vec[2]<- 99
print(my_vec)
```

c) Write a function called `my_func` that prints `This element is greater than 5!` if an element in `myvec` is greater than 5, and `This element is NOT greater than 5!` otherwise.
```{r}
greater_than_five <- function(Myvec) {
  for(element in Myvec) {
    if (element > 5) {
      print("this element is greater greater than 5!")}
    else {
      print("this element is NOT greater than 5!") } 
    } }
greater_than_five(my_vec)
```

## Task 3
a. Write a function named `get_fibonacci` that returns the n'th value in the [fibonacci sequence](https://www.mathsisfun.com/numbers/fibonacci-sequence.html). For example, `get_fibonacci(5)` should return 3; `get_fibonacci(7)` should return 8 and so on. ***Hint:** It is possible to call the same function from inside it. E.g., you can call `get_fibonacci()` from inside `get_fibonacci()`. Like inception!*
```{r}
get_fibonacci <- function(n) {
  if (n <= 0) {
    stop("N must be a positive integer")
  }
  if (n == 1)
    {return(0)}
  if (n==2) 
    {  return(1)}
  return(get_fibonacci(n-1)+ get_fibonacci(n-2))}

n <- 7
print(get_fibonacci(n))
```


b. Using your function in part (a), find and return the 19th value in the fibonacci sequence.
```{r}
get_fibonacci <- function(n) 
  { if (n <= 0) 
    { stop("N needs to be a positive integer")}
  if (n == 1)
    { return(0)}
  if (n==2) 
    {  return(1)}
  return(get_fibonacci(n-1)+ get_fibonacci(n-2))}

n <- 19
print(get_fibonacci(n))
```


<br>
You are done! 🏁 Don't forget to `commit` and `push` your .Rmd file to your Github clone repository before the due date.
<br>