---
title: "HW 2: Visualization, Base R"
subtitle: "Data 180, Professor Kennedy"
author: 
  name: "Sanskriti Rijal"
  email: "rijals@dickinson.edu"
date: '`r format(Sys.Date(), "%B %d, %Y")`' # no need to change date since this line grabs today's date
output: 
  html_document
---

Make sure that you have accepted the hw invitation link, and working with your clone repository.

**Due date:** September 28 beginning of class.

**Reminder:** You can insert R code in your Rmd file as a chunk in the following way:
```{r}
# Your code goes here
```
Make sure to insert a chunk for each (sub)section of the assignment.

**Theme of the assignment**: Let's practice visualization using base R and create cool maps. Load in the `Cars93` dataset by running the following code chunk:

```{r}
library(MASS)
data(Cars93)
```

The library function will load and attach add-on packages that have been previously installed on your machine. This will give you access to the functions and datasets available in the package. In general, you only need to install a package once (although it may require updating from time to time) but you need to access it using the library command every time you start a new session in R (and want the functionality). 

The questions below use the `Cars93` data frame as a source of data. This data frame is part of the MASS package and becomes available after loading it. You can get a view of the top of the data frame by using the head command. For more information, type `?Cars93`. Let's load `Cars93` in,
```{r}
head(Cars93)
```

# Base R
## Task 1
a) Create a Frequency table and Relative Frequency table (latter uses percentages) for the variable `Type`. Make sure to round all digits to 3 digits.
```{r} 
library(MASS)
data(Cars93)
F_table <- table(Cars93$Type)
rel_table <- prop.table(F_table) * 100
rel_table <- round(rel_table,3)
print(F_table)
print(rel_table)
```

b) Create a frequency bar graph for the variable `Type`. You can have count on the y axis.
```{r}
library(MASS)
library(ggplot2)
data(Cars93)

f_table <- table(Cars93$Type)

colors <- rainbow(length(f_table))
barplot(f_table, col=colors, names.arg= names(f_table), xlab = "Car type", ylab = "Count", main = "Frequence Bar Grpah for Car Types", cex.names = 1.2, cex.lab = 1.2, cex.axis=1.2)


```

*	Your axes should be labeled and legible
*	The bars should be of different colors
*	The plot should include a title using the main argument
*	The names, labels, and axes sizes (cex parameter) should be set to 1.2

c) Create a pie chart of the variable `Type`.
```{r}
library(MASS)
data(Cars93)
f_table <- table(Cars93$Type)
piechart <- pie(f_table, labels= paste(names(f_table), "(", f_table, ")", sep=""), main = "Pie Chart of Car Types", cex = 1.2)
print(piechart)

```

* The plot should include a title using the main argument.
*	The names, labels, and axes sizes (cex parameter) should be set to 1.2

d) What is the mode of the variable `Type`? Comment.
The mode is midsize with the frequency of 22.

## Task 2 
a) Create a histogram of the variable `Horsepower`.
```{r}
library(MASS)
data(Cars93)
histogram <- hist(Cars93$Horsepower, col="lightpink",xlab="Horsepower", ylab="Frequency", main= "Histogram of Horsepower", cex.axis=1.2, cex.lab=1.2, cex.names=1.2)
grid()
print(histogram)
```

*	Your axes should be labeled
*	The hist body should have a color (your choice, not too dark to obscure lines) 
*	The plot should include a title using the main argument
*	The names, labels, and axes sizes should be set to 1.2

b) Create a second histogram with breaks at 25 HP increments instead of 50.
```{r}
library(MASS)
data(Cars93)
breaks_ <- seq(0, max(Cars93$Horsepower)+25, by=25)
histogram <- hist(Cars93$Horsepower, breaks= breaks_, col="lightpink", xlab="Horsepower", ylab="Frequency", main="Histogram of Horsepower- 25 HP Increment", cex.axis = 1.2, cex.lab= 1.2, ces.names= 1.2)
grid()
print(histogram)
```

* Use the breaks argument to define the classes
* Other setting should be as in part (a)

c) Which histogram more closely follows or rule of thumb for the suggested number of classes? Explain.

The second histogram with 25 HP increments aligns more closely to the rule of thumb for the suggested number of classses as it has more bins bringing it closer to the recommended range and provides a better representation of the distribution. 



## Task 3 
a) What is the type of variable `Passengers`? (Categorical nominal/ordinal or Numerical discrete/continuous?) What happens if we were to make a histogram for `Passengers`? What is the problem with this graph?
The variable 'Passengers' is categorical nominal because each value represents the number of passengers a car model can seat. Since histograms show the distribution of numerical data, either discrete or continuous, the categorical nominal variable is not an appropriate choice because there is no numberical relationship between these categories. Instead, a bar graph or pie chart would be a better option to visualize the frequency.  

b) Correct the problem in part (a) by creating a more appropriate visualization. Make sure to label your axes, place a title, choose a custom color as in the previous graphs. (Bonus: use factor() with levels option to display x axis values with no observations on the plot.)
```{r}
library(MASS)
data(Cars93)
passengers <- factor(Cars93$Passengers, levels = 1:6)
passenger_count <-table(passengers)
colors <- "lightblue"
barplot(passenger_count, col=colors, xlab="Number of Passengers", ylab="Frequency", main="Bar Chart of Passenger Counts", names.arg=levels(passengers), cex.axis=1.2, cex.lab=1.2, cex.names=1.2)
grid()

```

## Task 4 
a) Make a histogram of the variable `Length` (the default breaks are fine). Other aspects of the plot should be as in part (a) in Task 2.
```{r}
library(MASS)
data(Cars93)
histogram_len <- hist(Cars93$Length, col="lightgreen", xlab="Length", ylab="Frequency", main="Histogram of Car Length", cex.axis=1.2, cex.lab= 1.2, cex.names=1.2)
grid()
print(histogram_len)
```

b) What is the shape of the distribution?
The shape of the distribution is aproximately bell shaped since it has a peak at the center and distribution of both sides of the peak. 

c) Find the mean and median of the variable `Length.` Do these answers meet your expectations from part (b)?
```{r}
library(MASS)
data(Cars93)
mean_len <- mean(Cars93$Length)
median_len <- median(Cars93$Length)
print(mean_len)
print(median_len)
```
Yes these answers meet my expetation as the mean is 183.2043 and the median is 183 which is roughly similar to each other and both are located at the center of the distribution.

You are done! ðŸ Don't forget to commit and push your .Rmd file to your Github repository before the due date.


